---
title: "5010 Project - Python to R"
author: "Project Team"
date: "3/23/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Contextual Inflation Analysis

Read file and drop the last row from the dataframe as we did not have this data in other tables to use this information
```{r wage}
df<- read.csv('Nina Charts (Wage, Gas).csv')
df_new<- head(df,-1)
```

## Data Cleaning - List of Years, Gas and House prices between years 1938 to 2009 and its equivalent in 2021

List of years,house prices,house price equivalent in 2021,minimum wage, minimum wage equivalent in 2021,gas prices from 1938 to 2009, 

```{r pressure, echo=FALSE}
years<- c(1938,1939,1945,1950,1956,1961,1963,1967,1968,1974,1975,1976,1978,
          1979,1980,1981,1989,1990,1991,1996,1997,2006,2007,2008,2009)

house_years<- c(3900,2938,7354,7354,11900,18000,20000,22900,23900,36200,39000,44300,55300,
                60600,63700,70400,123900,120000,130000,144100,145000,247900,226500,214300,326400)

house_2021 <- c(76503,54101,80333,80333,104346,151152,163163,179439,162141,181209,192977,204939,228099,228734,211075,
                208608,250714,229834,247882,234227,231889,306730,265457,256539,392358)

mw_year<- c(0.25,0.30,0.40,0.75,1.00,1.15,1.25,1.40,1.60,2.00,2.10,2.30,2.65,2.90,3.10,3.35,3.35,3.80,4.25,
            4.75,5.15,5.15,5.85,6.55,7.25)

mw_2021<-c(4.60,5.52,5.70,7.19,9.62,9.88,10.50,10.97,12.06,10.61,10.39,10.66,10.93,10.95,10.27,9.93,6.78,7.60,
           8.10,7.74,8.24,6.37,7.24,7.68,6.03)

gas_year<-c(0.20,0.19,0.27,0.29,0.31,0.31,0.30,0.32,0.35,0.39,0.53,0.57,0.62,0.63,0.86,1.19,1.00,1.14,1.11,1.23,
            1.23,2.80,2.80,3.27,2.35)
```


```{r}
df1<-data.frame(A = house_years,B=house_2021)
colnames(df1)<-c('Housing Cost','2021 Equivalent Cost')
rownames(df1)<-years
```


Formatting functions for ggplot  graph axis(source-below)
```{r}
#https://github.com/fdryan/R/blob/master/ggplot2_formatter.r
human_numbers <- function(x = NULL, smbl ="", signif = 1){
  humanity <- function(y){
    
    if (!is.na(y)){
      tn <- round(abs(y) / 1e12, signif)
      b <- round(abs(y) / 1e9, signif)
      m <- round(abs(y) / 1e6, signif)
      k <- round(abs(y) / 1e3, signif)
      
      if ( y >= 0 ){
        y_is_positive <- ""
      } else {
        y_is_positive <- "-"
      }
      
      if ( k < 1 ) {
        paste0( y_is_positive, smbl, round(abs(y), signif ))
      } else if ( m < 1){
        paste0 (y_is_positive, smbl,  k , "k")
      } else if (b < 1){
        paste0 (y_is_positive, smbl, m ,"m")
      }else if(tn < 1){
        paste0 (y_is_positive, smbl, b ,"bn")
      } else {
        paste0 (y_is_positive, smbl,  comma(tn), "tn")
      }
    } else if (is.na(y) | is.null(y)){
      "-"
    }
  }
  
  sapply(x,humanity)
}

#' Human versions of large currency numbers - extensible via smbl

human_gbp   <- function(x){human_numbers(x, smbl = "£")}
human_usd   <- function(x){human_numbers(x, smbl = "$")}
human_euro  <- function(x){human_numbers(x, smbl = "€")} 
human_num   <- function(x){human_numbers(x, smbl = "")}

```

# House Inflation Plot
```{r houseinflation}
library(ggplot2)

#HOuse Inflation Plot
ggplot(df1,aes(x=row.names(df1),group=1))+
  geom_line(aes(y =`Housing Cost`),color='grey')+
  geom_point(aes(y =`Housing Cost`),shape=21,color='black',fill='#69b3a2',size=2)+
  geom_line(aes(y =`2021 Equivalent Cost`),color='grey')+
  geom_point(aes(y =`2021 Equivalent Cost`),shape=21,color='black',fill='#728ab1',size=2)+
  ggtitle("Housing Inflation 1938-2021")+
  scale_y_continuous(labels=human_usd) +
  theme(axis.text.x = element_text(angle = 90))+
  xlab('Years')
  
```


## Minimum Wage vs Gas Price Inflation Plot

```{r}
df2<-data.frame(A = mw_year,B=mw_2021, C=gas_year )
colnames(df2)<-c('Minimum Wage','2021 Equivalent Wage', 'Price of Gas')
rownames(df2)<-years

ggplot(df2,aes(x=row.names(df2),group=1))+
  geom_line(aes(y =`Minimum Wage`),color='grey')+
  geom_point(aes(y =`Minimum Wage`),shape=21,color='black',fill='#b17e72',size=2)+
  geom_line(aes(y =`2021 Equivalent Wage`),color='grey')+
  geom_point(aes(y =`2021 Equivalent Wage`),shape=21,color='black',fill='#7274b1',size=2)+
  geom_line(aes(y =`Price of Gas`),color='grey')+
  geom_point(aes(y =`Price of Gas`),shape=21,color='black',fill='#b172a2',size=2)+
  ggtitle("Minimum Wage vs Gas Price Inflation Plot")+scale_y_continuous(labels=human_usd) +
  theme(axis.text.x = element_text(angle = 90))+
  xlab('Years')

```

## Weekly hours required to earn a living wage
list of US states including D.C
```{r}
states <- c('Alabama','Alaska','Arizona','Arkansas','California','Colorado','Conecticut',
          'Delaware','D.C.','Florida','Georgia','Hawaii','Idaho','Illinois','Indiana','Iowa',
          'Kansas','Kentucky','Louisiana','Maine','Maryland','Massachusetts','Michigan',
          'Minnesota','Mississippi','Missouri','Montana','Nebraska','Nevada','New Hampshire',
          'New Jersey','New Mexico','New York','North Carolina','North Dakota','Ohio',
          'Oklahoma','Oregon','Pennsylvania','Rhode Island','South Carolina','South Dakota',
          'Tennessee','Texas','Utah','Vermont','Virginia','Washington','West Virginia',
          'Wisconsin','Wyoming')
states
```


```{r}
myrange<-0:51
normal<-list(myrange)

lw_1a0c<-c(76,59,50,53,62,55,53,66,61,69,85,77,77,61,74,75,75,74,78,50,63,56,56,60,74,58,64,60,
         68,80,59,62,63,81,72,61,75,56,74,56,80,54,73,77,80,54,92,48,61,77,73)

lw_1a1c<-c(149,117,97,113,134,114,110,139,116,144,163,153,154,125,150,161,157,156,161,102,
         126,116,129,125,148,121,137,130,145,166,131,127,122,166,153,131,156,110,152,117,
         164,116,145,158,163,116,181,95,132,166,152)

lw_2a1w1c<-c(146,110,96,109,123,104,102,131,105,135,158,144,151,115,146,152,149,147,157,96,
           116,98,108,117,146,118,128,123,134,152,115,124,109,152,144,119,149,106,142,106,
           183,111,141,151,156,107,166,88,120,150,143)

lw_2aw1c_b<-c(82,64,53,62,73,62,60,76,62,78,89,83,86,68,83,88,86,86,89,56,68,62,70,68,81,67,76,
            71,80,90,71,70,66,91,84,72,86,60,83,64,90,64,80,86,90,63,98,52,71,91,84)
lw_2aw1c<-c(41,32,27,31,36,31,30,38,31,39,45,41,43,34,41,44,43,43,45,28,34,31,35,34,41,33,
          38,36,40,45,36,35,33,45,42,36,43,30,42,32,45,32,40,43,45,32,49,26,36,46,42)

ticks<-c(0,1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19,20,21,22,23,24,25,26,27,28,29,30,31,
       32,33,34,35,36,37,38,39,40,41,42,43,44,45,46,47,48,49,50)

```


```{r}
df3<-data.frame(A =lw_1a0c, B=lw_1a1c, C=lw_2a1w1c, D=lw_2aw1c )
colnames(df3)<-c('1 Adult, 0 Children','1 Adult, 1 Child', '2 Adults (1 Working), 1 Child','2 Working Adults, 1 Child')
rownames(df3)<-states
```


```{r}
ggplot(df3,aes(x=row.names(df3),group=1))+
  geom_line(aes(y =`1 Adult, 0 Children`),
            color='grey')+
  geom_point(aes(y =`1 Adult, 0 Children`),
             shape=21,color='black',
             fill='#b17e72',size=2)+
  geom_line(aes(y =`1 Adult, 1 Child`),
            color='grey')+
  geom_point(aes(y =`1 Adult, 1 Child`),
             shape=21,
             color='black',
             fill='#7274b1',
             size=2)+
  geom_line(aes(y =`2 Adults (1 Working), 1 Child`),
            color='grey')+
  geom_point(aes(y =`2 Adults (1 Working), 1 Child`),
             shape=21,
             color='black',
             fill='#b172a2',
             size=2)+
  geom_line(aes(y =`2 Working Adults, 1 Child`),
            color='grey')+
  geom_point(aes(y =`2 Working Adults, 1 Child`),
             shape=21,
             color='black',
             fill='#afb172',size=2)+
  ggtitle("Weekly Hours Required to Earn a Living Wage")+
  theme(axis.text.x = element_text(angle = 90))+
  xlab('US States')+ylab('Weekly Hours by Family Size')

```
## Analysis of Minimum Wage by US States
```{r}
df_state<-read.csv('Only State MW.csv')
df_state$Minimum.Wage<-sub("^.","",df_state$Minimum.Wage)
df_state$region <- tolower(df_state$State)
```

```{r}
states_map <-map_data("state")
```


```{r}
library(dplyr)
US_states<-left_join(states_map,df_state)
```

```{r}
centroids <- data.frame(region=tolower(state.name), long=state.center$x, lat=state.center$y)
centroids$abb<-state.abb[match(centroids$region,tolower(state.name))]
```


```{r}
p<-ggplot(data = US_states,mapping = aes(x=long,y=lat,
                                      group = group, 
                                      fill=as.numeric(Minimum.Wage)))+
  geom_polygon(color = "gray90", size = 0.1)+
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank(),
        axis.title.y=element_blank(),
        axis.title.x=element_blank(),
        title = element_text(size= 12,face="bold", 
                             colour = "black"),
                              plot.title = element_text(hjust=0.5),
                              strip.text = element_text(face="bold", size=9))+
  ggtitle('Minimum Wage by State')+
  scale_fill_distiller(name="Minimum Wage",palette = 'Spectral')

p2<- p+
  with(centroids,annotate(geom="text", x = long, y=lat, label = abb, 
                          size = 4,color="white",family="Times")
       )
p2
```

## US Counties map-> MW-Living Wage
```{r}
counties<-map_data('county')
df_counties = read.csv('Use This One v2 (For Real).csv')
```


```{r}
columns_list <- c('X1A_0C_LW','X1A_0C_PW','X1A_1C_LW','X2A_1W_1C_LW','X2AW_1C_LW','MW')
for (i in columns_list){
  df_counties[[i]]<-sub("^.","",df_counties[[i]])
  df_counties[[i]]<-as.numeric(df_counties[[i]])
}
```

```{r}
df_counties$X1A_0C_Diff = df_counties$MW - df_counties$X1A_0C_LW
df_counties$X1A_1C_Diff = df_counties$MW - df_counties$X1A_1C_LW
df_counties$X2A_1W_1C_Diff = df_counties$MW - df_counties$X2A_1W_1C_LW
df_counties$X2AW_1C_Diff = df_counties$MW - df_counties$X2AW_1C_LW
```


```{r}
library(maps)
library(dplyr)
library(tidyr)
library(tidyverse)

maps::county.fips %>%
  as.tibble %>% 
  extract(polyname, c("region", "subregion"), "^([^,]+),([^,]+)$") ->
  dfips

```

```{r}
map_data("county") %>% 
  left_join(dfips) ->
  dall
```

```{r}
dall$fips[dall$subregion == "okaloosa"] <- 12091
dall$fips[dall$subregion == "st martin"] <- 22099
dall$fips[dall$subregion == "currituck"] <- 37053
dall$fips[dall$subregion == "oglala dakota"] <- 
dall$fips[dall$subregion == "galveston"] <- 48167
dall$fips[dall$subregion == "accomack"] <- 51001
dall$fips[dall$subregion == "pierce" & dall$region == "washington" ] <- 53053
dall$fips[dall$subregion == "san juan" & dall$region == "washington" ] <- 53055
```




```{r}
dall %>% left_join(df_counties, by=c("fips"='Code'))-> US_counties
```

## Living Wage by County for 1 Adult 1Child Family size
```{r}
#Living Wage by County for 1 Adult 1Child Family size
p<-ggplot(data = US_counties,mapping = aes(x=long,y=lat,
                                         group = group, 
                                         fill=X1A_1C_LW))+
  geom_polygon(color = "gray90", size = 0.1)+
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank(),
        axis.title.y=element_blank(),
        axis.title.x=element_blank(),
        title = element_text(size= 12,face="bold", 
                             colour = "black"),
        plot.title = element_text(hjust=0.5),
        strip.text = element_text(face="bold", size=9))+
  ggtitle('Living Wage by County for 1 Adult 1Child Family size')+
  scale_fill_distiller(name="Living Wage",palette = 'Spectral',labels=scales::dollar_format(prefix = '$'))
p
```
## Gap between Minimum Wage and Living Wage by County for 1 Adult 1Child Family size
```{r}
#Gap between Minimum Wage and Living Wage by County for 1 Adult 1Child Family size
p<-ggplot(data = US_counties,mapping = aes(x=long,y=lat,
                                           group = group, 
                                           fill=X1A_1C_Diff))+
  geom_polygon(color = "gray90", size = 0.1)+
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank(),
        axis.title.y=element_blank(),
        axis.title.x=element_blank(),
        title = element_text(size= 12,face="bold", 
                             colour = "black"),
        plot.title = element_text(hjust=0.5),
        strip.text = element_text(face="bold", size=9))+
  ggtitle('             Gap between Minimum Wage and Living Wage for 1 Adult 1Child Family size by County')+
  scale_fill_continuous(name="Gap b/w MW and LW",type = 'viridis',labels=scales::dollar_format(prefix = '$'))
p
```
## Living Wage by County for 1 Adult 0Child Family size
```{r}
#Living Wage by County for 1 Adult 0Child Family size
p<-ggplot(data = US_counties,mapping = aes(x=long,y=lat,
                                           group = group, 
                                           fill=X1A_0C_LW))+
  geom_polygon(color = "gray90", size = 0.1)+
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank(),
        axis.title.y=element_blank(),
        axis.title.x=element_blank(),
        title = element_text(size= 12,face="bold", 
                             colour = "black"),
        plot.title = element_text(hjust=0.5),
        strip.text = element_text(face="bold", size=9))+
  ggtitle('Living Wage by County for 1 Adult 0Child Family size')+
  scale_fill_distiller(name="Living Wage",palette = 'Spectral',labels=scales::dollar_format(prefix = '$'))
p
```

```{r}
#Gap between Minimum Wage and Living Wage by County for 1 Adult 0Child Family size
library("scales")
p<-ggplot(data = US_counties,mapping = aes(x=long,y=lat,
                                           group = group, 
                                           fill=X1A_0C_Diff))+
  geom_polygon(color = "gray55", size = 0.1)+
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank(),
        axis.title.y=element_blank(),
        axis.title.x=element_blank(),
        title = element_text(size= 12,face="bold", 
                             colour = "black"),
        plot.title = element_text(hjust=0.5),
        strip.text = element_text(face="bold", size=9))+
  ggtitle('                     Gap b/w Minimum Wage-Living Wage for Single Workers by County')+
  scale_fill_continuous(name="MW and LW Gap",type = 'viridis',labels=scales::dollar_format(prefix = '$'))
p

```

## Living Wage by County for 2 Adult(1 working) 1Child Family size

```{r}
p<-ggplot(data = US_counties,mapping = aes(x=long,y=lat,
                                           group = group, 
                                           fill=X2A_1W_1C_LW))+
  geom_polygon(color = "gray90", size = 0.1)+
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank(),
        axis.title.y=element_blank(),
        axis.title.x=element_blank(),
        title = element_text(size= 12,face="bold", 
                             colour = "black"),
        plot.title = element_text(hjust=0.5),
        strip.text = element_text(face="bold", size=9))+
  ggtitle('Living Wage by County for 2 Adult(1 working) 1Child Family size')+
  scale_fill_distiller(name="Living Wage",palette = 'Spectral',labels=scales::dollar_format(prefix = '$'))
p

```

## Gap between Minimum Wage and Living Wage by County for 2 Adult(1 working) 1Child Family size
```{r}

p<-ggplot(data = US_counties,mapping = aes(x=long,y=lat,
                                           group = group, 
                                           fill=X2A_1W_1C_Diff))+
  geom_polygon(color = "gray90", size = 0.1)+
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank(),
        axis.title.y=element_blank(),
        axis.title.x=element_blank(),
        title = element_text(size= 12,face="bold", 
                             colour = "black"),
        plot.title = element_text(hjust=0.5),
        strip.text = element_text(face="bold", size=9))+
  ggtitle('   Gap between Minimum Wage and Living Wage by County for 2 Adult(1 working) 1Child Family size')+
  scale_fill_continuous(name="Gap b/w MW and LW",type = 'viridis',labels=scales::dollar_format(prefix = '$'))
p
```

## Living Wage by County for 2 Working Adult 1Child Family size
```{r}
#Living Wage by County for 2 Working Adult 1Child Family size
p<-ggplot(data = US_counties,mapping = aes(x=long,y=lat,
                                           group = group, 
                                           fill=X2AW_1C_LW))+
  geom_polygon(color = "gray90", size = 0.1)+
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank(),
        axis.title.y=element_blank(),
        axis.title.x=element_blank(),
        title = element_text(size= 12,face="bold", 
                             colour = "black"),
        plot.title = element_text(hjust=0.5),
        strip.text = element_text(face="bold", size=9))+
  ggtitle('Living Wage by County for 2 Working Adult 1Child Family size')+
  scale_fill_distiller(name="Living Wage",palette = 'Spectral',labels=scales::dollar_format(prefix = '$'))
p
```


```{r}
#Gap between Minimum Wage and Living Wage by County for 2 Working Adult 1Child Family size
p<-ggplot(data = US_counties,mapping = aes(x=long,y=lat,
                                           group = group, 
                                           fill=X2A_1W_1C_Diff))+
  geom_polygon(color = "gray90", size = 0.1)+
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank(),
        axis.title.y=element_blank(),
        axis.title.x=element_blank(),
        title = element_text(size= 12,face="bold", 
                             colour = "black"),
        plot.title = element_text(hjust=0.5),
        strip.text = element_text(face="bold", size=9))+
  ggtitle('Gap between Minimum Wage and Living Wage by County for 2 Working Adult 1Child Family size')+
  scale_fill_continuous(name="Gap b/w MW and LW",type= 'viridis',labels=scales::dollar_format(prefix = '$'))
p
```


```{r}
bls_data <- read.csv('BLS Chlorpeth Data.csv')

dall %>% left_join(bls_data, by=c("fips"='Area.Code'))-> bls_average


p<-ggplot(data = bls_average,mapping = aes(x=long,y=lat,
                                           group = group, 
                                           fill=Extrapolated.Average.Hourly.Wage))+
  geom_polygon(color = "gray90", size = 0.1)+
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank(),
        axis.title.y=element_blank(),
        axis.title.x=element_blank(),
        title = element_text(size= 12,face="bold", 
                             colour = "black"),
        plot.title = element_text(hjust=0.5),
        strip.text = element_text(face="bold", size=9))+
  ggtitle('BLS 2020 Hourly Wage')+
  scale_fill_continuous(name="Average Hourly Wage",type = 'gradient',labels=scales::dollar_format(prefix = '$'))
p
```

