---
title: "Research Report"
author: "Carolyn Fiore, Shruti Biradar, and Kevin Russell"
date: "4/25/2022"
output: pdf_document
knit: (function(inputFile, encoding) { 
      out_f <- paste0(tools::file_path_sans_ext(basename(inputFile)), ".pdf");
      rmarkdown::render(inputFile,
                        encoding=encoding,
                        output_file=file.path(dirname(inputFile), out_f)) })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(modelr)
library(foreach)
library(iterators)
library(glmnet)
```

# Initialization

## Functions

```{r}
# Custom function to apply parallel processing
parallel_proc <- function(num_of_clusters = -1){
  library(parallel)
  library(doParallel)
  # https://cran.r-project.org/web/packages/doParallel/index.html
  
  if (num_of_clusters == -1) {
    core_count <- parallel::detectCores() - 1
  }
  
  for (proc in ps::ps_children()) {
    if (ps::ps_name(proc) == "Rscript.exe") {
      ps::ps_kill(proc)
    }
  }
  
  sock_cluster <- makePSOCKcluster(core_count)
  registerDoParallel(sock_cluster)
  return(sock_cluster)
}

# Custom function to plot PCA graph
geom_biplot <- function(model,
                        pc_x = 'PC1',
                        pc_y = 'PC2',
                        x_coords = c(),
                        y_coords = c(),
                        point_color = 'black',
                        point_alpha = 1,
                        param_color = 'black',
                        arrow_color = 'blue',
                        arrow_alpha = 0.75,
                        arrow_size = 0.1,
                        arrow = TRUE,
                        circle = FALSE,
                        circle_type = 't',
                        circle_level = 0.95,
                        circle_color = 'grey70',
                        var_name_text_size = 3,
                        unit_len = FALSE,
                        unit_len_vector_amp = 1,
                        arrow_len_vector_amp = 0.6,
                        centroids = FALSE) {
  require(ggplot2)
  require(tidyverse)
  
  serr_func <- function(z) {
    sd(z) / sqrt(length(z))
  }
  
  model.data <- as_tibble(model$x)
  model.rotation <- tibble(varnames = row.names(model$rotation),
                            as_tibble(model$rotation))
  
  if (arrow) {
    if (unit_len) {
      model.rotation <- model.rotation |>
        mutate(vector1 = unit_len_vector_amp * get(pc_x),
               vector2 = unit_len_vector_amp * get(pc_y))
    } else {
      model.mult <- min((max(model.data[, pc_y]) -
                           min(model.data[, pc_y]) /
                           (max(model.rotation[, pc_y]) -
                              min(model.rotation[, pc_y]))),
                        (max(model.data[, pc_x]) -
                           min(model.data[, pc_x]) /
                           (max(model.rotation[, pc_x]) -
                              min(model.rotation[, pc_x]))))
      model.rotation <- model.rotation |>
        mutate(vector1 = arrow_len_vector_amp * model.mult * (get(pc_x)),
               vector2 = arrow_len_vector_amp * model.mult * (get(pc_y)))
    }
  }
  
  if (typeof(point_color) == 'character') {
    model.plot <- model.data |>
      ggplot(aes_string(x = pc_x, y = pc_y)) +
      geom_point(color = point_color,
                 alpha = point_alpha)
  } else {
    model.plot <- model.data |>
      ggplot(aes_string(x = pc_x, y = pc_y)) +
      geom_point(aes(color = point_color),
                 alpha = point_alpha)
  }
  
  if (arrow) {
    model.plot <- model.plot +
      geom_hline(aes(yintercept = 0),
                 size = arrow_size) +
      geom_vline(aes(xintercept = 0),
                 size = arrow_size)
    
    if (unit_len & (length(x_coords) < 1 | length(y_coords) < 1)) {
      x_coords <- c(-1, 1)
      y_coords <- c(-1, 1)
      
      if (circle) {
        model.angle <- seq(-pi, pi, length = 50)
        model.circle <- tibble(x = sin(model.angle),
                               y = cos(model.angle))
      }
    }
  } else if (typeof(point_color) != 'character') {
    centroids = TRUE
  }
  
  if (centroids) {
    gg_centroids <- aggregate(cbind(get(pc_x),
                                    get(pc_y)) ~
                                point_color,
                              model.data, mean)
    names(gg_centroids) <- c('point_color', pc_x, pc_y)
    
    se <- aggregate(cbind(se.x = get(pc_x),
                          se.y = get(pc_y)) ~ point_color,
                    model.data,
                    serr_func)
    gg_centroids <- merge(gg_centroids,
                          se,
                          by = "point_color")
    
    
    model.plot <- model.plot +
        geom_point(data = gg_centroids,
                   size = 2,
                   aes(color = point_color)) +
        geom_errorbar(data = gg_centroids,
                      aes(ymin = PC2 - se.y,
                          ymax = PC2 + se.y,
                          color = point_color),
                      width = 0.1) +
        geom_errorbarh(data = gg_centroids,
                       aes(xmin = PC1 - se.x,
                           xmax = PC1 + se.x,
                           color = point_color),
                       height=0.1)
  }
  
  if (length(x_coords) < 1 | length(y_coords) < 1) {
    model.plot <- model.plot +
      coord_equal()
  } else {
    model.plot <- model.plot +
      coord_cartesian(xlim = x_coords,
                               ylim = y_coords)
  }
  
  if (arrow) {
    model.plot <- model.plot +
      geom_text(data = model.rotation,
                aes(x = vector1,
                    y = vector2,
                    label = varnames),
                size = var_name_text_size,
                vjust = 1,
                color = param_color) +
      geom_segment(data = model.rotation,
                   aes(x = 0,
                       y = 0,
                       xend = vector1,
                       yend = vector2),
                   arrow = arrow(length = unit(arrow_size, 'cm')),
                   alpha = arrow_alpha,
                   color = arrow_color)
  }
  
  if (circle & unit_len) {
    model.plot <- model.plot +
      geom_path(aes(x, y),
              data = model.circle,
              colour = circle_color)
  } else if (circle) {
    model.plot <- model.plot +
      stat_ellipse(type = circle_type,
                   level = circle_level,
                   color = circle_color)
  }
  
  return(model.plot)
}
```

This is a function to graph principal component analysis graphs and for loading parallel processing.

## Data Load

```{r}
# File Path and Directory
script_dir <- dirname(rstudioapi::getSourceEditorContext()$path)
pov_fp <- paste(script_dir,
                 'percentage_magic.csv',
                 sep = '/')
cpi_gdp_fp <- paste(script_dir,
                 'cpi_gdp_data.csv',
                 sep = '/')

# Read data
cpi_gdp_data <- read_csv(cpi_gdp_fp,
                         show_col_types = FALSE)
pov_data <- read_csv(pov_fp,
                     show_col_types = FALSE) |>
  unique() |>
  inner_join(cpi_gdp_data, by = 'fips')

# Remove dupes and calculate average ethnicity poverty percent
tidy_data <- pov_data |>
  select(!c(long, lat, ID)) |>
  unique() |>
  mutate(eth_pov_pct = (`other%pov` +
                  `aia%pov` +
                  `as%pov` +
                  `hl%pov` +
                  `hp%pov` +
                  `tr%pov` +
                  `aa%pov` +
                  `wnh%pov`) / 8)

# Train/Test
set.seed(1)
pov_part <- resample_partition(tidy_data,
                                p = c(train=0.8,
                                      test=0.2))
train_data <- as_tibble(pov_part$train)
test_data <- as_tibble(pov_part$test)
```

Loading and combining data with train test split of 80% train and 20% test.

## Pre-Processing

```{r}
# NA's to zeros
train_data[is.na(train_data)] <- 0

# Drop all columns that has same value for all rows
train_data <- train_data |>
  purrr::keep(~length(unique(.x)) != 1)
```

We are amputating by replacement of NAs to zeros. Also, dropping columns with same values for all rows.

# Classification

## Finding Classification

```{r}
ggplot(train_data,
       aes(x = pov_pct,
           y = eth_pov_pct)) +
    geom_point(color = 'pink') +
    geom_smooth(formula = y ~ s(x, bs = "cs"),
                method = 'gam',
                color = 'black') +
    geom_smooth(formula = y ~ x,
                method='lm',
                color='red') +
    labs(x = "Poverty Percent",
         y = "Average Ethnicity Poverty Percent",
         title = 'EDA - Poverty Percent cutoff might be 12.5%') +
    theme(plot.title = element_text(hjust = 0.5))
```

Poverty Percent shows a cutoff of potentially 12.5% where Average Ethnicity Poverty Percent starts to increase. It is unclear whether 12.5% is acceptable, but lets confirm this with the test data set

## Confirming Classification

```{r}
ggplot(test_data,
       aes(x = pov_pct,
           y = eth_pov_pct)) +
    geom_point(color = 'green') +
    geom_smooth(formula = y ~ s(x, bs = "cs"),
                method = 'gam',
                color = 'black') +
    geom_smooth(formula = y ~ x,
                method='lm',
                color='red') +
    labs(x = "Poverty Percent",
         y = "Average Ethnicity Poverty Percent",
         title = 'CDA - Poverty Percent cutoff at 12.5%') +
    theme(plot.title = element_text(hjust = 0.5))
```

Test data set confirms that 12.5% and higher should be classified as Impoverished, and anything below 12.5% should be classified as Not Impoverished. Let's classify this data set and remove the bias.

## Creating Classification 

```{r}
# To convert government category to integer
gov_parties <- unique(tidy_data$government)
gov_parties <- tibble(gov_party_id = 1:length(gov_parties),
                      gov_party_name = gov_parties)

# Classify, Convert to Int, and Remove Bias
tidy_data <- tidy_data |>
  left_join(gov_parties, by = c('government' = 'gov_party_name')) |>
  mutate(poverty = factor(ifelse(pov_pct >= 12.5,
                          'Impoverished',
                          'Not Impoverished'),
                          levels = c('Impoverished',
                          'Not Impoverished'))) |>
  select(!c(ends_with('_belowPoverty'),
            ends_with('_TE'),
            ends_with('%pov'),
            pov_pct,
            eth_pov_pct,
            government))

# Train/Test
set.seed(2)
pov_part <- resample_partition(tidy_data,
                               p = c(train=0.8,
                                     test=0.2))
train_data <- as_tibble(pov_part$train) |>
  select(!c(region.x, subregion, st_fips_code))
test_data <- as_tibble(pov_part$test) |>
  select(!c(region.x, subregion, st_fips_code))
```

## Pre-Processing Again

```{r}
# NA's to zeros
train_data[is.na(train_data)] <- 0
test_data[is.na(test_data)] <- 0

# Drop all columns that has same value for all rows
train_data <- train_data |>
  purrr::keep(~length(unique(.x)) != 1)
test_data <- test_data |>
  purrr::keep(~length(unique(.x)) != 1)
```

# Exploratory Data Analysis

## Principal Component Fitting

```{r}
# Separate Poverty measurement
pca_class_ind <- train_data |>
  dplyr::select(poverty) |>
  as.vector()

# Re-code Poverty to integer and fit PCA comp
pca_fit_cmp <- train_data |>
  dplyr::select(- poverty) |>
  prcomp(center = TRUE,
         scale. = TRUE)

# Calculate Cumulative Proportion and find 80% point
cum_porp <- cumsum(pca_fit_cmp$sdev^2 / sum(pca_fit_cmp$sdev^2))
perc_80_x <- which(abs(cum_porp - .80) == min(abs(cum_porp - .80)))

plot(cum_porp,
     type="b",
     main = '80% of Variation is by 2 Components',
     xlab = 'Principal Components by #',
     ylab = 'Cumulative Proportion')

points(perc_80_x,
       cum_porp[perc_80_x],
       col = "red",
       cex = 2,
       pch = 20)
```

By fitting Principal Component Computation, the first two Principal Components explain 80% of the pattern variation. This is great because PCA can reduce the dimension from 73 candidate predictors to 2 candidate predictors.

## Principal Component Graph

```{r}
geom_biplot(pca_fit_cmp,
            circle = TRUE,
            unit_len = TRUE,
            unit_len_vector_amp = 3.5,
            point_color = pca_class_ind$poverty,
            param_color = 'purple',
            arrow_color = 'brown',
            centroids = TRUE) +
  labs(x = "PC1",
       y = "PC2",
       color = 'Poverty',
       title = 'EDA - Household Items Lead to Poverty?') +
  theme(plot.title = element_text(hjust = 0.5))
```

Impoverished is strongly correlated with fips, X1A_0C_Diff and potentially a lot of CPI categories in the upper left part of the graph. Let's re-graph by zooming into the left quadrant.

```{r}
geom_biplot(pca_fit_cmp,
            circle = FALSE,
            unit_len = TRUE,
            unit_len_vector_amp = 3.4,
            point_color = pca_class_ind$poverty,
            param_color = 'purple',
            arrow_color = 'brown',
            centroids = TRUE,
            x_coords = c(-.425, 0),
            y_coords = c(0, 1)) +
  labs(x = "PC1",
       y = "PC2",
       color = 'Poverty',
       title = 'EDA - Household Items Lead to Poverty?') +
  theme(plot.title = element_text(hjust = 0.5))
```

After looking at the quadrant, there are 14 CPI measurements that potentially correlate with Impoverished. Those measurements are listed here:

* CPI_Commodities_Less_Food
* CPI_Commodities_Less_Food_And_Beverages
* CPI_Education_And_Communication
* CPI_Apparel
* CPI_Recreation
* CPI_Household_Furnishings_And_Operations
* CPI_Other_Personal_Services
* CPI_Education_And_Communication_Services
* CPI_Transportation_Commodities_Less_Motor_Fuel
* CPI_Other_Goods
* CPI_New_And_Used_Motor_Vehicles
* CPI_Durables
* CPI_Household_Furnishings_And_Supplies
* CPI_Education_And_Communication_Commodities

Let's examine the distributions for each of these measurements along with X1A_0C_Diff and fips against poverty in groups of 5-6 distributions.

## Candidate Predictors Distributions

```{r}
g1 <- ggplot(train_data,
       aes(x = X1A_0C_Diff,
           y = poverty)) +
  geom_boxplot(color = 'green') +
  labs(x = "X1A 0C Difference",
       y = "Poverty",
       title = 'Impoverished Increased') +
  theme(plot.title = element_text(hjust = 0.5))

g2 <- ggplot(train_data,
       aes(x = fips,
           y = poverty)) +
  geom_boxplot(color = 'blue') +
  labs(x = "Fips",
       y = "Poverty",
       title = 'Impoverished Decreased') +
  theme(plot.title = element_text(hjust = 0.5))

g3 <- ggplot(train_data,
       aes(x = CPI_Commodities_Less_Food,
           y = poverty)) +
  geom_boxplot(color = 'brown') +
  labs(x = "Commodities Less Food",
       y = "Poverty",
       title = 'Impoverished Decreased') +
  theme(plot.title = element_text(hjust = 0.5))

g4 <- ggplot(train_data,
       aes(x = CPI_Commodities_Less_Food_And_Beverages,
           y = poverty)) +
  geom_boxplot(color = 'orange') +
  labs(x = "Commodities Less Food and Beverages",
       y = "Poverty",
       title = 'Impoverished Decreased') +
  theme(plot.title = element_text(hjust = 0.5))

g5 <- ggplot(train_data,
       aes(x = CPI_Education_And_Communication,
           y = poverty)) +
  geom_boxplot(color = 'purple') +
  labs(x = "Education and Communication",
       y = "Poverty",
       title = 'Impoverished Neutral') +
  theme(plot.title = element_text(hjust = 0.5))

gridExtra::grid.arrange(g1, g2, g3, g4, g5,
                        top = 'Candidate Predictors - Group One')
```

All measurements median influence Impoverished except for Education and Communication. However, X1A 0C Difference and Fips has the biggest change in distribution. Let's look at the second grouping of distributions.

```{r}
g1 <- ggplot(train_data,
       aes(x = CPI_Apparel,
           y = poverty)) +
  geom_boxplot(color = 'green') +
  labs(x = "Apparel",
       y = "Poverty",
       title = 'Impoverished Neutral') +
  theme(plot.title = element_text(hjust = 0.5))

g2 <- ggplot(train_data,
       aes(x = CPI_Recreation,
           y = poverty)) +
  geom_boxplot(color = 'blue') +
  labs(x = "Recreation",
       y = "Poverty",
       title = 'Impoverished Increased') +
  theme(plot.title = element_text(hjust = 0.5))

g3 <- ggplot(train_data,
       aes(x = CPI_Household_Furnishings_And_Operations,
           y = poverty)) +
  geom_boxplot(color = 'brown') +
  labs(x = "Household Furnishings And Operations",
       y = "Poverty",
       title = 'Impoverished Neutral') +
  theme(plot.title = element_text(hjust = 0.5))

g4 <- ggplot(train_data,
       aes(x = CPI_Other_Personal_Services,
           y = poverty)) +
  geom_boxplot(color = 'orange') +
  labs(x = "Other Personal Services",
       y = "Poverty",
       title = 'Impoverished Decreased') +
  theme(plot.title = element_text(hjust = 0.5))

g5 <- ggplot(train_data,
       aes(x = CPI_Education_And_Communication_Services,
           y = poverty)) +
  geom_boxplot(color = 'purple') +
  labs(x = "Education And Communication Services",
       y = "Poverty",
       title = 'Impoverished Increased') +
  theme(plot.title = element_text(hjust = 0.5))

gridExtra::grid.arrange(g1, g2, g3, g4, g5,
                        top = 'Candidate Predictors - Group Two')
```

Household Furnishings And Operations and Apparel means are neutral against Impoverished whereas the other three measurements affected Impoverished in some way. It seems that Recreation and Other Personal Services had the biggest affect against poverty. Let's look at the third grouping of distributions.

```{r}
g1 <- ggplot(train_data,
       aes(x = CPI_Transportation_Commodities_Less_Motor_Fuel,
           y = poverty)) +
  geom_boxplot(color = 'green') +
  labs(x = "Transportation Commodities Less Motor Fuel",
       y = "Poverty",
       title = 'Impoverished Neutral') +
  theme(plot.title = element_text(hjust = 0.5))

g2 <- ggplot(train_data,
       aes(x = CPI_Other_Goods,
           y = poverty)) +
  geom_boxplot(color = 'blue') +
  labs(x = "Other Goods",
       y = "Poverty",
       title = 'Impoverished Neutral') +
  theme(plot.title = element_text(hjust = 0.5))

g3 <- ggplot(train_data,
       aes(x = CPI_New_And_Used_Motor_Vehicles,
           y = poverty)) +
  geom_boxplot(color = 'brown') +
  labs(x = "New And Used Motor Vehicles",
       y = "Poverty",
       title = 'Impoverished Neutral') +
  theme(plot.title = element_text(hjust = 0.5))

g4 <- ggplot(train_data,
       aes(x = CPI_Durables,
           y = poverty)) +
  geom_boxplot(color = 'orange') +
  labs(x = "Durables",
       y = "Poverty",
       title = 'Impoverished Neutral') +
  theme(plot.title = element_text(hjust = 0.5))

g5 <- ggplot(train_data,
       aes(x = CPI_Household_Furnishings_And_Supplies,
           y = poverty)) +
  geom_boxplot(color = 'purple') +
  labs(x = "Household Furnishings And Supplies",
       y = "Poverty",
       title = 'Impoverished Decreased') +
  theme(plot.title = element_text(hjust = 0.5))

g6 <- ggplot(train_data,
       aes(x = CPI_Education_And_Communication_Commodities,
           y = poverty)) +
  geom_boxplot(color = 'purple') +
  labs(x = "Education And Communication Commodities",
       y = "Poverty",
       title = 'Impoverished Decreased') +
  theme(plot.title = element_text(hjust = 0.5))

gridExtra::grid.arrange(g1, g2, g3, g4, g5, g6,
                        top = 'Candidate Predictors - Group Three')
```

Every measurement's median is neutral except for Household Furnishings And Supplies and Education And Communication Commodities. However, Education And Communication Commodities had the biggest change for poverty.

Overall, X1A 0C Difference, Fips, Recreation, Other Personal Services, Household Furnishings And Supplies, and Education And Communication Commodities had the biggest impact against poverty.

For the Null Hypothesis, X1A 0C Difference, Fips, Recreation, Other Personal Services, Household Furnishings And Supplies, and Education And Communication Commodities will not impact poverty classification. The alternative hypothesis is that at least two of these measurements will impact poverty classification. The alpha will be 0.01.
